Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-04-29 paper/article_jos.qmd
V2.1

---
title: "Improving real-time trend estimates using local parametrisation of polynomial regression filters"
format:
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    keep-tex: true
    toc: false
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
    cite-method: biblatex
crossref:
  tbl-prefix: table
  fig-prefix: figure
  sec-prefix: section
  eq-prefix: equation

@@

:::: {.abstract}
This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
In particular, this theory allows us to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates (for the last estimates): first including a timeliness criterion to minimise the phase shift;
second, with procedure for parametrising asymmetric filters locally While they are generally parametrised globally, which can be suboptimal around turning points.
An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
The paper also shows how these results can be easily reproduced using the R package `rjd3filters` dedicated to the manipulation of moving averages.

@@

Analysis of the economic cycle, and in particular the early detection of turning points in a series, is a major topic in the analysis of economic outlook.
To this end, economic indicators are generally seasonally adjusted.
However, in order to improve their interpretability, it may be necessary to perform additional smoothing in order to reduce noise, and thus analyse the trend-cycle component.
By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series.

A moving average, or linear filter, is a statistical method that consists of applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
For example, the X-11 decomposition method [@x11], used in X-13ARIMA-SEATS software [@x13],  uses several moving averages to estimate the main components of a time series. 
Symmetric filters are applied to the center of the series, but when it comes to estimate the most recent points, all of these methods must rely on asymmetric filters. 
For trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
For example, for trend-cycle extraction, X-11 uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
Since the predicted values are linear combinations of past values, this amounts to applying asymmetric averages at the end of the series.

@@

Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
In particular, it becomes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to two drawbacks of those methods.
Firstly,  adding, in the optimisation process, a timeliness criterin to directly control the phase shift (i.e. the delay in detecting turning points).
Secondly, calibrating asymmetric filter locally, while they are generally calibrated globally  using thus assumptions that may be wrong locally.
Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and can therefore be easily used.
All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).

@@

The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and a an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e. when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

- When $p=f$ the moving average is said to be *centered*. 
If we also have $\forall k:\:\theta_{-k} = \theta_k$, the moving average $M_{\boldsymbol\theta}$ is said to be *symmetric*. 
In this case, the quantity $h=p=f$ is called the *bandwidth*.

@@

Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
Noting $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e. movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component.

@Fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e. when no point in the future is known).
The gain function is greater than 1 on the frequencies associated with the trend-cycle ($[0, 2\pi/12]$): this means that the trend-cycle is well preserved and that short cycles of 1 to 2 years ($[2\pi/24,2\pi/12]$) are even amplified.
At other frequencies, the gain function is less than 1, but always positive: this means that the series smoothed by this moving average will always contain noise, even if it is attenuated.
Analysis of the phase shift shows that the shorter the cycles, the greater the phase shift: this means that on series smoothed by this moving average, turning points could be detected at the wrong date.

@@

Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg). 
In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.

@@

2. Apply symmetric filters to extended forecast series.
This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).

@@

In this paper we will focus on @proietti2008 approach. 
@Sec-sym-lp and @sec-lppasymf describes their approach to build symmetric moving averages (used for the final estimates of the trend-cycle component) and asymmetric moving averages (used for real-time estimates).
This paper proposed two extensions of their approach to build asymetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parameterisation of a parameter usually parameterised globally (@sec-lptimeliness), which is used in the simulation.

@@

Using @proietti2008's notation, we assume that our time series $y_t$ can be decomposed into
$$
y_t=\mu_t+\varepsilon_t,
$$
where $\mu_t$ is the trend and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
Within a neighbourhood $h$ of $t$, the local trend $\mu_t$ is approximated by a polynomial of degree $d$, such that $\mu_t\simeq m_{t}$ with:
$$
\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
$$
The problem of trend extraction is equivalent to estimating $m_t=\beta_0$ (the constant in the previous formula).

@@

- the number of neighbors $2h+1$ (or the window $h$): if it is too small, then there will be too little data for the estimates (resulting in high variance in the estimates), and if it is too large, the polynomial approximation will probably be wrong, leading to biased estimates.

@@

Rewriting @eq-lpp:
$$
\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
$$ {#eq-lpgeneralmodel}
where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X.$
Defining the variance error as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ (to directly take into account the weights associated to the observations) and with $[\boldsymbol U,\boldsymbol Z] = \boldsymbol X$ we thus find the model associated to symmetric filters.
The objective is to find an asymmetric filter $\boldsymbol v$ which uses $q$ points in the future and minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain (polynomial) constraints.
The matrices $\boldsymbol y$, $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned as follows:
$$
\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
$$
where $\boldsymbol y_p$ is a matrix $(h+q+1)\times 1$ which denotes the set of available observations (from $y_{t-h}$ to $y_{t+q}$) and $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned in the same way.
The constraints are defined as $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$: they represents the polynomial preservation constraints we want to impose on the asymmetric filter.
@proietti2008 show that the problem is equivalent to finding $v$ which minimises:
$$
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
$$ {#eq-lppasym}
where $\boldsymbol l$ is a vector of Lagrange multipliers.

@@

- If the ratio is large ($3.5< R$) then a 23 term filter is used (to remove more noise).

- If the ratio is small ($R<1$) a 9 term filter is used.

@@

To measure the phase shift between the input series and the filtered one, @eq-lppasym introduces a criterion, $T_g$, called *timeliness*.
When a moving average $\boldsymbol v$ is applied, the level input signal is also altered by the gain function. 
That’s why the timeliness criterion depends on the gain and phase shift functions ($\rho_{\boldsymbol v}$ and $\varphi_{\boldsymbol v}$), the link between both functions being made by a penalty function $f$.
The criterion measures phase shift for the frequencies in the interval $[\omega_1, \omega_2]$:
$$
T_g(\boldsymbol v)=\int_{\omega_{1}}^{\omega_{2}}f(\rho_{\boldsymbol v}(\omega),\varphi_{\boldsymbol v}(\omega))\ud\omega
$$
As mentioned in @sec-propMM, in the case of trend-cycle extraction we can set $\omega_1=0$ and for monthly series $\omega_2=2\pi/12.$
For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show thant in that case the criteria can be written in a quadratic form:
$$
T_g(\boldsymbol v) = \transp{\boldsymbol v}\boldsymbol T\boldsymbol v
$$
with $T\boldsymbol=\int_{\omega_{1}}^{\omega_{2}}\boldsymbol{Si}(\omega)\transp{\boldsymbol{Si}(\omega)}\ud\omega$ and $\boldsymbol{Si}(\omega)=(\sin(p\omega),\dots, \sin (f\omega))$ if $\boldsymbol v$ uses $p$ points in the past and $f$ points in the future.

@@

One drawback is that $T_g(\boldsymbol v)$ is not normalized: the weight $\alpha_T$ has no economic sense.
It is then difficult to calibrate the value of $\alpha_T$ since the value of $\lvert\delta/\sigma\rvert$ must also be defined.
Hence, in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parametrisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).

@@

The turning points are clearly detected by the local estimators (changement in the slope), when $|\delta_1/\sigma|$ is estimated in the real-time estimate.
However, the real-time local estimates of $|\delta_1/\sigma|$ are noisy: the varability of this estimates, used to build asymmetric filters, will then lead to another source of revisions in the intermediate estimates.

@@

There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series rjd3filters`.

@@

The different methods are compared on simulated and real data. 
For all series, a symmetric 13-term filter is used: we then assume than the bandwidth is fixed, as in the seasonal adjustement method X-11.
A nearest neighbour bandwidth is also tested (13-term asymmetric filters: real-time estimates uses observation between $y_{t-12}$ and $y_{t}$, the second asymmetric filters uses obversations between $y_{t-11}$ and $y_{t+1}$, etc.) but with no improvement in the results (see supplemental material).
More complexe methods to locally select bandwidth, like the one proposed by @fan1992variable, were not tested.
All the methods to create asymmetric filters are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
The ARIMA model is determined automatically, using @autoarima algorithm (function `forecast::auto.arima()`) on the last 12 years, with no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).

@@

- High signal-to-noise ratio (i.e. low I-C ratio and low variability): $\sigma_e^2=0.2$ and $\rho = 3.0, 3.5$ or $4.0$ (I-C ratio between 0.9 and 0.7);

- Medium signal-to-noise ratio (i.e. medium I-C ratio and medium variability): $\sigma_e^2=0.3$ and $\rho = 1.5,\, 2.0$ or $3.0$ (I-C ratio between 2.3 and 1.4);

- Low signal-to-noise ratio (i.e. high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).

@@

The differences between the methods are also illustrated using an example from the FRED-MD database (@fredmd) containing economic series on the United States.
The series studied correspond to the database published in November 2022.
The series studied is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
This point and this series were chosen because the economic turnaround is particularly visible in the raw series (@fig-ce16ov-previmp-lp), because this series is used for dating economic cycles and because the FRED-MD database facilitates the reproducibility of the results (thanks to the availability of series published on past dates).
Studying the series up to January 2020, this series has a medium variability: a symmetric filter of 13 terms is therefore appropriate.

>>>>

---
title: "Improving real-time trend estimates using local parametrisation of polynomial regression filters"
format:
  docx:
    toc: false
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    keep-tex: true
    toc: false
    fig-pos: 'H'
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
    cite-method: biblatex
crossref:
  tbl-prefix: table
  fig-prefix: figure
  sec-prefix: section
  eq-prefix: equation

@@

:::: {.abstract}
This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
It enables to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates: first including a timeliness criterion to minimise the phase shift;
second, with procedure for parametrising asymmetric filters locally while they are generally parametrised globally, which can be suboptimal around turning points.
An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
The results are reproducible and all the methods can be easily applied using the R package `rjd3filters`.

@@

Analysis of the economic cycle, and in particular the early detection of turning points, is a major topic in the analysis of economic outlook.
To achieve this, economic indicators are generally seasonally adjusted.
However, in order to improve their interpretability, it may be necessary to perform additional smoothing to reduce noise, thereby to analyse the trend-cycle component.
By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series. 
Thus, even though trend-cycle extraction methods are generally applied to seasonally adjusted series, the estimation of these series also depends on trend-cycle estimation methods.

A moving average, or linear filter, is a statistical method that consists of applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
They are ubiquitous in business cycle extraction and seasonal adjustment methods.
For example, the X-11 seasonal adjustment method [@ladiray2011seasonal], used in X-13ARIMA-SEATS software [@x13], uses several moving averages to estimate the main components of a time series. 
Symmetric filters are applied to the centre of the series and for trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
For example, for trend-cycle extraction, X-11 uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
Since the predicted values are linear combinations of past values, this amounts to applying asymmetric moving averages at the end of the series.

@@

Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
In particular, this class encompass Henderson's symmetric filter and Musgrave's asymmetric filters.
The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods.
Firstly,  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
Secondly, calibrating asymmetric filter locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).

@@

The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], a usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

- When $p=f$ the moving average is said to be *centred*. 
If we also have $\forall k:\:\theta_{-k} = \theta_k$, the moving average $M_{\boldsymbol\theta}$ is said to be *symmetric*. 
In this case, the quantity $h=p=f$ is called the *bandwidth*.

@@

Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
Noting $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e., movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component.

@Fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e., when no point in the future is known).
The gain function is greater than 1 on the frequencies associated with the trend-cycle ($[0, 2\pi/12]$): this means that the trend-cycle is well preserved and that short cycles of 1 to 2 years ($[2\pi/24,2\pi/12]$) are even amplified.
At other frequencies, the gain function is less than 1, but always positive: this means that the series smoothed by this moving average will always contain noise, even if it is attenuated.
Analysis of the phase shift shows that the shorter the cycles, the greater the phase shift: this means that on series smoothed by this moving average, turning points could be detected at the wrong date.

@@

\justifying
*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified of 10% ($G_{\boldsymbol\theta}(\pi/6)=1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*\
*Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg).* 
*In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.*

@@

2. Apply symmetric filters to extended forecast series.
This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extended by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).

@@

In this paper we will focus on @proietti2008 approach. 
@Sec-sym-lp and @sec-lppasymf describes their approach to build symmetric moving averages (used for the final estimates of the trend-cycle component) and asymmetric moving averages (used for real-time estimates).
This paper proposed two extensions of their approach to build asymmetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parameterisation of a parameter usually parameterised globally (@sec-lptimeliness), which is used in the simulation.

@@

Using @proietti2008's notation, we assume that our time series $y_t$ can be decomposed into
$$
y_t=TC_t+\varepsilon_t,
$$
where $TC_t$ is the trend-cycle and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
Within a neighbourhood $h$ of $t$, the local trend $\mu_t$ is approximated by a polynomial of degree $d$, such that $\mu_t\simeq m_{t}$ with:
$$
\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
$$
The problem of trend extraction is equivalent to estimating $m_t=\beta_0$ (the constant in the previous formula).

@@

- the number of neighbours $2h+1$ (or the window $h$): if it is too small, then there will be too little data for the estimates (resulting in high variance in the estimates), and if it is too large, the polynomial approximation will probably be wrong, leading to biased estimates.

@@

Rewriting @eq-lpp:
$$
\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
$$ {#eq-lpgeneralmodel}
where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X.$
Defining the variance error as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ (to directly take into account the weights associated to the observations) and with $[\boldsymbol U,\boldsymbol Z] = \boldsymbol X$ we thus find the model associated to symmetric filters.
The objective is to find an asymmetric filter $\boldsymbol v$ which uses $q$ points in the future and minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain (polynomial) constraints.
The matrices $\boldsymbol y$, $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned as follows:
$$
\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
$$
where $\boldsymbol y_p$ is a matrix $(h+q+1)\times 1$ which denotes the set of available observations (from $y_{t-h}$ to $y_{t+q}$) and $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned in the same way.
The constraints are defined as $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$: they represent the polynomial preservation constraints we want to impose on the asymmetric filter.
$\boldsymbol Z$ is then associated to the polynomial constraints of the symmetric filter not imposed on asymmetric filter.
@proietti2008 show that the problem is equivalent to finding $v$ which minimises:
$$
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
$$ {#eq-lppasym}
where $\boldsymbol l$ is a vector of Lagrange multipliers.

@@

- If the ratio is large ($3.5< R$) then a 23-term filter is used (to remove more noise).

- If the ratio is small ($R<1$) a 9-term filter is used.

@@

To measure the phase shift between the input series and the filtered one, @eq-lppasym introduces a criterion, $T_g$, called *timeliness*.
When a moving average $\boldsymbol v$ is applied, the level input signal is also altered by the gain function. 
That’s why the timeliness criterion depends on the gain and phase shift functions ($\rho_{\boldsymbol v}$ and $\varphi_{\boldsymbol v}$), the link between both functions being made by a penalty function $f$.
The criterion measures phase shift for the frequencies in the interval $[\omega_1, \omega_2]$:
$$
T_g(\boldsymbol v)=\int_{\omega_{1}}^{\omega_{2}}f(\rho_{\boldsymbol v}(\omega),\varphi_{\boldsymbol v}(\omega))\ud\omega
$$
As mentioned in @sec-propMM, in the case of trend-cycle extraction we can set $\omega_1=0$ and for monthly series $\omega_2=2\pi/12.$
For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show than in that case the criteria can be written in a quadratic form:
$$
T_g(\boldsymbol v) = \transp{\boldsymbol v}\boldsymbol T\boldsymbol v
$$
with $T\boldsymbol=\int_{\omega_{1}}^{\omega_{2}}\boldsymbol{Si}(\omega)\transp{\boldsymbol{Si}(\omega)}\ud\omega$ and $\boldsymbol{Si}(\omega)=(\sin(p\omega),\dots, \sin (f\omega))$ if $\boldsymbol v$ uses $p$ points in the past and $f$ points in the future.

@@

One drawback is that $T_g(\boldsymbol v)$ is not normalized: the weight $\alpha_T$ has no economic sense.
It is then difficult to calibrate the value of $\alpha_T$ since the value of $\lvert\delta/\sigma\rvert$ must also be defined.
Hence, in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
However, in future studies, we could imagine a two-step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parametrisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).

@@

The turning points are clearly detected by the local estimators (slope tends towards zero) when $|\delta_1/\sigma|$ is estimated in the real-time.
However, the real-time local estimates of $|\delta_1/\sigma|$ are noisy: the variability of these estimates, used to build asymmetric filters, will then lead to another source of revisions in the intermediate estimates.

@@

There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series using `rjd3filters`.

@@

The different methods are compared on simulated and real data. 
For all series, a symmetric 13-term filter is used: we then assume than the bandwidth is fixed, as in the seasonal adjustment method X-11.
A nearest neighbour bandwidth is also tested (13-term asymmetric filters: real-time filter uses observations between $y_{t-12}$ and $y_{t}$, the second asymmetric filters uses observations between $y_{t-11}$ and $y_{t+1}$, etc.) but with no improvement in the results (see supplemental material).
More complexe methods to locally select bandwidth, like the one proposed by @fan1992variable, were not tested.
All the methods to create asymmetric filters are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
The ARIMA model is determined automatically, using @autoarima algorithm (function `forecast::auto.arima()`) on the last 12 years, with no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).

@@

- High signal-to-noise ratio (i.e., low I-C ratio and low variability): $\sigma_e^2=0.2$ and $\rho = 3.0, 3.5$ or $4.0$ (I-C ratio between 0.9 and 0.7);

- Medium signal-to-noise ratio (i.e., medium I-C ratio and medium variability): $\sigma_e^2=0.3$ and $\rho = 1.5,\, 2.0$ or $3.0$ (I-C ratio between 2.3 and 1.4);

- Low signal-to-noise ratio (i.e., high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).

@@

The differences between the methods are also illustrated using an example from the FRED-MD database (@fredmd) containing economic series on the United States. 
This database facilitates the reproducibility of the results thanks to the availability of series published on past dates.
The series studied correspond to the database published in November 2022.
The series studied is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
This turning point was chosen because it is particularly visible in the raw series (@fig-ce16ov-previmp-lp) and this series was select because it is used for dating economic cycles.
Studying the series up to January 2020, this series has a medium variability: a symmetric filter of 13 terms is therefore appropriate.
